---
title: "Wine Sale Forecasting"
output: md_document
date: "2024-10-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Load required Libraries
```{r}
library(openxlsx)  
library(timetk)  
library(dplyr)  
library(lubridate)  
library(recipes)  
library(rsample)  
library(parsnip)  
library(workflows)  
library(forecastML)  
library(corrplot)  
library(doParallel)  
library(timeDate)  
library(astsa)  
library(ggplot2)  
library(forecast)  
library(bsts)  
library(nlme)  
library(splines) 
library(caret)
```


## 2 Set Parallel Processing
```{r}
# setting up parallel processing
registerDoSEQ()
cl = makePSOCKcluster(4)
registerDoParallel(cl)
```

## 3 Function for Performance Metrics
```{r}
Metrics.fun = function(x,y) {
  AE= abs(x-y)  # Absolute difference
  AE.mean = mean(AE)  # Mean AE
  SE = AE^2  # Squared AE
  MSE = mean(SE)  # Mean of squared error
  RMSE = sqrt(MSE) # Root of MSE
  RMSE.se = sd(SE) # Standard error of SE
  Metrics = data.frame(MAE = AE.mean,
                       RMSE=RMSE)
  Metrics
}
```


## 4 Load Dataset
```{r}
setwd("C:\\Users\\YCHEN\\Git\\Weather on wine sales")

load("Weekly wine sales data.RData")
str(ApothicUnit)

```

## 5 Exploratory Data Analysis (EDA)
```{r}
ApothicUnit %>%
  plot_time_series(DATE,apothicPremiumUnit,
                   .interactive=F,
                   .facet_scales="free",
                   .color_var=year(DATE),
                   .facet_ncol=2,
                   .x_lab="Date",
                   .y_lab="Apothic (Unit)")+
  labs(title = "Fig. 1 Weekly Wine Sales")
```
Note the uptrending of sales over the years.

```{r}
ApothicUnit %>%
  dplyr::group_by(Year) %>%
  plot_time_series(DATE,apothicPremiumUnit,
                   .interactive=F,
                   .facet_scales="free",
                   .facet_ncol=2,
                   .x_lab="Date",
                   .y_lab="Apothic (Unit)")+
  labs(title = "Fig. 2 Weekly Wine Sales by Year")
```
Monthly variation of sales is obvious.

```{r}
corrplot(cor(ApothicUnit[,c(2:8,10:16,18:25)]),
         method="square",order="hclust",
         title = "Fig. 3 Correlation Plot/Matrix of All Variables",
         mar = c(1,3,1,1))
```
Note the extremely high correlation among many related variables.


Select a subset of variables for prediction.
```{r} 
ApothicUnit_df = ApothicUnit %>%
  dplyr::select(apothicPremiumUnit,PRCP.mean, TMIN.mean,AWND.mean,
                SNOW.mean,SNWD.mean,Year, Month, WeekOfYear,WeekNo) 
names(ApothicUnit_df)

corrplot(cor(ApothicUnit_df[,2:6]),method="number",order="hclust",
         title = "Fig. 4 Correlation Plot/Matrix of Selected Variables",
         mar = c(1,3,1,1))
```

Check NAs and zero variance or near-zero variance
```{r}
apply(is.na(ApothicUnit_df),2,sum)
nearZeroVar(ApothicUnit_df,names=T)
```


Plots of wine sales against weather variables
```{r}
ggplot(ApothicUnit_df,aes(x=PRCP.mean,y=apothicPremiumUnit))+
  geom_point()+
  geom_smooth(method="lm",se=F)+
  geom_smooth(method="loess",se=F,color="red")
ggplot(ApothicUnit_df,aes(x=TMIN.mean,y=apothicPremiumUnit))+
  geom_point()+
  geom_smooth(method="lm",se=F)+
  geom_smooth(method="loess",se=F,color="red")
ggplot(ApothicUnit_df,aes(x=AWND.mean,y=apothicPremiumUnit))+
  geom_point()+
  geom_smooth(method="lm",se=F)+
  geom_smooth(method="loess",se=F,color="red")
ggplot(ApothicUnit_df,aes(x=SNOW.mean,y=apothicPremiumUnit))+
  geom_point()+
  geom_smooth(method="lm",se=F)+
  geom_smooth(method="loess",se=F,color="red")
ggplot(ApothicUnit_df,aes(x=SNWD.mean,y=apothicPremiumUnit))+
  geom_point()+
  geom_smooth(method="lm",se=F)+
  geom_smooth(method="loess",se=F,color="red")
```

Check time series autocorrelation.
```{r}
lag2.plot(as.ts(ApothicUnit_df$apothicPremiumUnit),
          as.ts(ApothicUnit_df$PRCP.mean),
          max.lag=15)
lag2.plot(as.ts(ApothicUnit_df$apothicPremiumUnit),
          as.ts(ApothicUnit_df$TMIN.mean),
          max.lag=15)
lag2.plot(as.ts(ApothicUnit_df$apothicPremiumUnit),
          as.ts(ApothicUnit_df$AWND.mean),
          max.lag=15)
lag2.plot(as.ts(ApothicUnit_df$apothicPremiumUnit),
          as.ts(ApothicUnit_df$SNOW.mean),
          max.lag=15)
lag2.plot(as.ts(ApothicUnit_df$apothicPremiumUnit),
          as.ts(ApothicUnit_df$SNWD.mean),
          max.lag=15)
```


## 6 Create Dummy Holiday Variables
```{r}
ApothicUnit_df$week.end = seq(as.Date("2012-01-15"),as.Date("2017-12-31"),by="1 week")
ApothicUnit_df$week.start = ApothicUnit_df$week.end - 7
```

Christmas
```{r}
Christmas.Dates = timeDate::ChristmasDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$Christmas[i] =  ifelse(length(base::intersect(week,as.Date(Christmas.Dates))) == 0,0,1)
}
```

Thanksgiving
```{r}
ThanksgivingDay.Dates = timeDate::USThanksgivingDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$ThanksgivingDay[i] = ifelse(length(base::intersect(week,as.Date(ThanksgivingDay.Dates))) == 0,0,1)
}
```

New Year's day
```{r}
NewYearsDay.Dates = timeDate::NewYearsDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$NewYearsDay[i] = ifelse(length(base::intersect(week,as.Date(NewYearsDay.Dates))) == 0,0,1)
}

```
Independence day
```{r}
IndependenceDay.Dates = timeDate::USIndependenceDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$IndependenceDay[i] = ifelse(length(base::intersect(week,as.Date(IndependenceDay.Dates))) == 0,0,1)
}
```

Labor day
```{r}
LaborDay.Dates = timeDate::USLaborDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$LaborDay[i] = ifelse(length(base::intersect(week,as.Date(LaborDay.Dates))) == 0,0,1)
}
```

Memorial day
```{r}
MemorialDay.Dates = timeDate::USMemorialDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$MemorialDay[i] = ifelse(length(base::intersect(week,as.Date(MemorialDay.Dates))) == 0,0,1)
}
```

Presidents day
```{r}
PresidentsDay.Dates = timeDate::USPresidentsDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$PresidentsDay[i] = ifelse(length(base::intersect(week,as.Date(PresidentsDay.Dates))) == 0,0,1)
}
```

MLKing'S day
```{r}
MLKingsBirthday.Dates = timeDate::USMLKingsBirthday(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$MLKingsBirthday[i] = ifelse(length(base::intersect(week,as.Date(MLKingsBirthday.Dates))) == 0,0,1)
}
```

Veterans day
```{r}
VeteransDay.Dates = timeDate::USVeteransDay(min(ApothicUnit_df$Year):max(ApothicUnit_df$Year))
for (i in 1:nrow(ApothicUnit_df)) {
  week = seq(ApothicUnit_df[i,]$week.start+1,ApothicUnit_df[i,]$week.end,length.out=7)
  ApothicUnit_df$VeteransDay[i] = ifelse(length(base::intersect(week,as.Date(VeteransDay.Dates))) == 0,0,1)
}
```

## 7 Train and Test Data Split
271 weeks for training and 40 weeks for testing.
```{r}
ApothicUnit_train = ApothicUnit_df[1:(nrow(ApothicUnit_df)-40),]
ApothicUnit_test = ApothicUnit_df[(nrow(ApothicUnit_df)-40+1):nrow(ApothicUnit_df),]
names(ApothicUnit_train)

TS_Forecast_train_df = ApothicUnit_train %>%
  dplyr::select(1:6,11) %>%
  dplyr::rename("Apothic" = "apothicPremiumUnit")
str(TS_Forecast_train_df)

TS_Forecast_test_df = ApothicUnit_test %>%
  dplyr::select(1,11) %>%
  dplyr::rename("Apothic" = "apothicPremiumUnit")
str(TS_Forecast_test_df)
```

## 8 Model Building and Testing
### 8.1 SARIMA models
logarithm transformation does not help too much.
```{r}
plot.ts(ts(TS_Forecast_train_df[,1]))
acf2(ts(TS_Forecast_train_df[,1]))

plot.ts(log10(ts(TS_Forecast_train_df[,1])))  # seems better
acf2(log10(ts(TS_Forecast_train_df[,1])))  # seems better
```

#### 8.1.1 SARIMA models without exogenous variables
Autocorrelation strong after 15 lags.
```{r}
lag1.plot(ts(TS_Forecast_train_df[,1]),max.lag=15)
```
Perioditiy of 5 weeks?
```{r}
spectrum(ts(TS_Forecast_train_df[,1]),log="no")
spectrum(ts(TS_Forecast_train_df[,1]),log="no",plot=F)
```

Autodetection of arima model (p,d,q = 0,1,2)
```{r}
auto.arima(TS_Forecast_train_df[,1],trace=T,ic="aicc",seasonal.test="hegy") 
acf2(TS_Forecast_train_df[,1])  # ar3 is adequate?
```

Fitting model with p,d,q = 3,1,0
```{r}
sarima(ts(TS_Forecast_train_df[,1]),3,1,0)   
sarima_model1_pred = sarima.for(ts(TS_Forecast_train_df[,1]),40,3,1,0)$pred 
Metrics.fun(sarima_model1_pred,TS_Forecast_test_df[,1])
```

Fitting model with p,d,q = 0,1,2
```{r}
sarima(ts(TS_Forecast_train_df[,1]),0,1,2) # best model selected by auto.arima
sarima_model2_pred = sarima.for(ts(TS_Forecast_train_df[,1]),40,0,1,2)$pred # 
Metrics.fun(sarima_model2_pred,TS_Forecast_test_df[,1])
```

Fitting model with  p,d,q, P,D,Q,S = 3,1,0,1,0,0,52 (the best so far) 
```{r}
sarima_model3 = sarima(ts(TS_Forecast_train_df[,1]),3,1,0,1,0,0,S=52)
sarima_model3_pred = sarima.for(ts(TS_Forecast_train_df[,1]),40,3,1,0,1,0,0,S=52)$pred
Metrics.fun(sarima_model3_pred,TS_Forecast_test_df[,1])
```

#### 8.1.2 SARIMA models with exogenous variables
One pro of using exogenous variables is that exogenous variables may significantly increase forecasting accuracy.
One con of using exogenous variables is that all the exogenous variables for the forecasting periods need to be forecasted first.


##### 8.1.2.1 Forecasting five weather variables for 40 weeks
```{r}
names(TS_Forecast_train_df)
names(TS_Forecast_test_df)
```

PRCP.mean
```{r}
plot.ts(ts(TS_Forecast_train_df[,2]))
auto.arima(ts(TS_Forecast_train_df[,2]))
sarima(ts(TS_Forecast_train_df[,2]),2,0,2,1,0,0,S=52)
PRCP.sarima.S = sarima.for(ts(TS_Forecast_train_df[,2]),40,2,0,2,1,0,0,S=52)
```

TMIN.mean
```{r}
plot.ts(ts(TS_Forecast_train_df[,3]))
auto.arima(ts(TS_Forecast_train_df[,3]))
sarima(ts(TS_Forecast_train_df[,3]),1,0,1,1,0,0,S=52)
TMIN.sarima.S = sarima.for(ts(TS_Forecast_train_df[,3]),40,1,0,1,1,0,0,S=52)
```


AWND.mean
```{r}
plot.ts(ts(TS_Forecast_train_df[,4]))
auto.arima(ts(TS_Forecast_train_df[,4]))
sarima(ts(TS_Forecast_train_df[,4]),1,0,1,1,0,0,S=52)
AWND.sarima.S = sarima.for(ts(TS_Forecast_train_df[,4]),40,1,0,1,1,0,0,S=52)
```


SNOW.mean
```{r}
plot.ts(ts(TS_Forecast_train_df[,5]))
auto.arima(ts(TS_Forecast_train_df[,5]))
sarima(ts(TS_Forecast_train_df[,5]),1,0,3,1,0,0,S=52)
SNOW.sarima.S = sarima.for(ts(TS_Forecast_train_df[,5]),40,1,0,3,1,0,0,S=52)
```

SNWD.mean
```{r}
plot.ts(ts(TS_Forecast_train_df[,6]))
auto.arima(ts(TS_Forecast_train_df[,6]))
sarima(ts(TS_Forecast_train_df[,6]),1,0,5,1,0,0,S=52)
SNWD.sarima.S = sarima.for(ts(TS_Forecast_train_df[,6]),40,1,0,5,1,0,0,S=52)
```


New data for forecasting
```{r}
names(ApothicUnit_train)
names(ApothicUnit_test)

newxreg = data.frame(PRCP.mean = PRCP.sarima.S$pred,
                     TMIN.mean=TMIN.sarima.S$pred,
                     AWND.mean=AWND.sarima.S$pred,
                     SNOW.mean=SNOW.sarima.S$pred,
                     SNWD.mean=SNWD.sarima.S$pred,
                     Christmas=ApothicUnit_test$Christmas,
                     NewYearsDay=ApothicUnit_test$NewYearsDay,
                     IndependenceDay=ApothicUnit_test$IndependenceDay,
                     LaborDay=ApothicUnit_test$LaborDay,
                     ThanksgivingDay=ApothicUnit_test$ThanksgivingDay,
                     MemorialDay=ApothicUnit_test$MemorialDay,
                     PresidentsDay=ApothicUnit_test$PresidentsDay,
                     MLKingsBirthday=ApothicUnit_test$MLKingsBirthday,
                     VeteransDay=ApothicUnit_test$VeteransDay,
                     WeekOfYear = ApothicUnit_test$WeekOfYear,
                     WeekNo=ApothicUnit_test$WeekNo,
                     Month=as.factor(ApothicUnit_test$Month),
                     Year=ApothicUnit_test$Year)
newxreg = newxreg %>%
  dplyr::mutate(SNOW.Binary = ifelse(newxreg$SNOW.mean>=0.2,1,0),
                SNWD.Binary = ifelse(newxreg$SNWD.mean>=2,1,0)) %>%
  dplyr::mutate(SNOW.Binary.Back1 = c(SNOW.Binary[-1],0),
                SNWD.Binary.Back1 = c(SNWD.Binary[-1],0)) %>%
  as.data.frame()
```

str(newxreg)


##### 8.1.2.2 Five weather variables as exogenous
Model
```{r}
sarima_model3_exo = sarima(ts(TS_Forecast_train_df[,1]),0,1,2,1,0,0,S=52,
                       xreg=TS_Forecast_train_df[,2:6])
acf2(sarima_model3_exo$fit$residuals,max.lag=100)
auto.arima(sarima_model3_exo$fit$residuals)
```

Prediction
```{r}
sarima_model3_exo_pred = sarima.for(ts(TS_Forecast_train_df[,1]),40,0,1,2,1,0,0,S=52,
                                xreg=TS_Forecast_train_df[,2:6],
                                newxreg=newxreg[,1:5])
Metrics.fun(sarima_model3_exo_pred$pred,TS_Forecast_test_df[,1])  
```
Note
(1) The performance of the model with five weather variables is worse than that with any exogenous variables.
(2) Adding holidays does not improve model performance (data not shown).



### 8.2 bsts (Bayesian Structural Time Series) models
#### 8.2.1 Local level model with S=52
```{r}
ss = AddLocalLevel(list(),TS_Forecast_train_df[,1]) 
ss = AddSeasonal(ss,TS_Forecast_train_df[,1],nseasons=52)
bsts.Local.Level = bsts(TS_Forecast_train_df[,1],
                        state.specification=ss,niter=1000)
bsts.Local.Level.Pred = predict(bsts.Local.Level,horizon=40,burn=200)
plot(bsts.Local.Level.Pred)
Metrics.fun(bsts.Local.Level.Pred$mean,TS_Forecast_test_df[,1])
```

